---
title: "PS2 - Econometrics I"
author: "Gustavo Henrique & Bruno Tonholo"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: sandstone
    toc: TRUE
    toc_float: TRUE
---

# Question 1
In the conditional logit model, the market share of a good $j \in \{0, 1, 2\}$ (or the  probability of choosing the good $j$) is given by:

$$ s_j = \mathbb{P}[\delta_j+\epsilon_j>\max\limits_{k \neq j}\{\delta_k+\epsilon_k\}] = \frac{e^{\delta_j}}{\sum_{k=0}^2 e^{\delta_k}} $$

Then, since $\delta_0=0$, the market shares for each good are:

$$ s_0 = \frac{1}{1+e^{\delta_1}+e^{\delta_2}}, \qquad s_1 = \frac{e^{\delta_1}}{1+e^{\delta_1}+e^{\delta_2}}, \qquad s_2 = \frac{e^{\delta_2}}{1+e^{\delta_1}+e^{\delta_2}} $$

So, dividing $s_1$ by $s_0$ gives us:

$$ \frac{s_1}{s_0}=\frac{\frac{e^{\delta_1}}{1+e^{\delta_1}+e^{\delta_2}}}{\frac{1}{1+e^{\delta_1}+e^{\delta_2}}}=e^{\delta_1} $$
Therefore, by taking the natural logarithm of both sides, we obtain:

$$ \delta_1=\log(\frac{s_1}{s_0})=\log(s_1)-\log(s_0) $$

as desired.

# Question 2
Consider a student $i$ who is going to start a PhD in economics at EPGE. He then needs to buy a new laptop to help him with his studies, but is unsure which one to choose. His utility in obtaining a laptop can be represented as follows:

$$ U_i(x)=\beta_0+\beta_1p_x+\beta_2model_x+\beta_3opsys_x+\beta_4strg_x+\epsilon_{ix} $$

where $x$ is the amount of laptop, $p_x$ is the price of laptop, $opsys_x$ represents the operating system, $strg_x$ is the storage capacity and $\epsilon_{ix}$ is the random error term, which follows an Extreme Value (EV) distribution.

# Question 3
Encoding:
```{r}
# Loading needed packages
pacman::p_load(
  tidyverse,
  here
)

# Importing database
blp_data = readRDS(here("data", "blp-data.Rds"))

# Calculating the non-participants market share per year
mrkt_share_non_participants = blp_data %>%
  group_by(year) %>%
  summarize(s_0 = 1 - sum(share, na.rm = TRUE))

# Adding s_0 values to the original database
blp_data = blp_data %>%
  left_join(mrkt_share_non_participants, by = "year") %>% 
  # Calculating the δ_j values
  mutate(delta_j = log(share) - log(s_0))

# Printing result
head(blp_data %>%
  # Selecting convenient columns
  select(model_id, share, s_0, delta_j))
```

# Question 4
Encoding:
```{r}
# Estimating
model = lm(delta_j ~ hpwt + air + mpd + space + price, data = blp_data)

# Printing the result
summary_model = summary(model)
summary_model
```

# Question 5
Encoding:
```{r}
# Extracting analytical OLS standard errors
ols_se = summary_model$coefficients[, "Std. Error"]

# Implementing non-parametric bootstrap
set.seed(123)  # Set seed for reproducibility
B = 100       # Number of bootstrap repetitions
# Creating a matrix to store bootstrap coefficients
bootstrap_estimates = matrix(NA, nrow = B, ncol = length(coef(model)))  

for (b in 1:B) {
  # Resampling data indices with replacement
  sample_indices = sample(1:nrow(blp_data), replace = TRUE)
  bootstrap_sample = blp_data[sample_indices, ]
  
  # Fitting the model to the bootstrap sample
  bootstrap_model = lm(delta_j ~ hpwt + air + mpd + space + price, data = bootstrap_sample)
  
  # Storing the estimated coefficients
  bootstrap_estimates[b, ] = coef(bootstrap_model)
}

# Calculating bootstrap standard errors
bootstrap_se = apply(bootstrap_estimates, 2, sd)

# Creating a comparison table for OLS and bootstrap standard errors
comparison = data.frame(
  attribute = names(ols_se),
  ols_se = ols_se,
  bootstrap_se = bootstrap_se
) %>% 
  mutate(attribute = ifelse(attribute == "(Intercept)", "intercept", attribute))

# Printing the comparison results
comparison
```
