---
title: "Problem Set 1 - Econometrics I (Part 2)"
author: "Gustavo Henrique & Bruno Tonholo"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: sandstone
    toc: TRUE
    toc_float: TRUE
---

# Question 4: Panel Data
## Item (a)
```{r}
## Loading needed packages
pacman::p_load(
  tidyverse,
  plm,
  fixest
)

## Importing database
psid_df = read.csv("WAGES_PSID_7yrs.csv")

## Cleaning database
psid_df = psid_df %>% 
  mutate(exp_sqrd = exp^2) %>% 
  rename("id" = "ID",
         "year" = "YEAR") %>% 
  select(-c(3:7, 9:10))

## Implementing the within transformation
group_means = psid_df %>%
  group_by(id) %>%
  summarise(
    lwage_mean = mean(lwage),
    across(c(edu, ms, exp, exp_sqrd), mean, .names = "{.col}_mean"),
    .groups = "drop"
  )
within_df = left_join(psid_df, group_means, by = "id") %>% 
  mutate(lwage_dem = lwage - lwage_mean,
         edu_dem = edu - edu_mean,
         exp_dem = exp - exp_mean,
         exp_sqrd_dem = exp_sqrd - exp_sqrd_mean,
         ms_dem = ms - ms_mean)

## Estimating a fixed effects "within" estimator
within_est = lm(lwage_dem ~ -1 + edu_dem + ms_dem + exp_dem + exp_sqrd_dem,
           data = within_df)

## Estimating the same model using plm package as benchmark
plm_fe = plm(lwage ~ edu + ms + exp + exp_sqrd,
           data = psid_df,
           model = "within",
           effect = "individual",
           index = c("id", "year"))

## Extracting results
coef_within = coef(summary(within_est))
coef_fe = coef(summary(plm_fe))

## Creating a comparison DataFrame
comparison_fe = data.frame(
  Estimate_within = coef_within[, "Estimate"],
  Estimate_plm_fe = coef_fe[, "Estimate"],
  StdError_within = as.double(sprintf("%.10f", coef_within[, "Std. Error"])),
  StdError_plm_fe = coef_fe[, "Std. Error"]
)

## Summarizing results
comparison_fe
```
The results show an inverse relationship between wage and marital status, a direct relationship between wage and experience and a very small inverse relationship between wage and squared experience. In terms of statistical significance, both experience and squared experience are strong, while marital status fail to reject the null at 5% level.
Also, by comparing with the benchmark estimation using *plm*, there is no difference between the coefficients, and there is negligible difference in terms of standard error, p-value and t-statistic. Notice that, given the structure of the dataset, the coefficient for the education variable was suppressed, since it has no intra-individual variation.

## Item (b)
```{r}
## Estimating a random effects model
# Extracting relevant variables
id = psid_df$id
time = psid_df$year
y = psid_df$lwage
X = as.matrix(psid_df[, c("edu", "ms", "exp", "exp_sqrd")])
colnames(X) = c("edu", "ms", "exp", "exp_sqrd")

# Computing overall means
y_bar = mean(y)
X_bar = colMeans(X)

# Extracting group means for variables
y_i_bar = within_df$lwage_mean
X_i_bar = within_df %>%
  select(ends_with("_mean")) %>% 
  select(-lwage_mean) %>% 
  as.matrix()

# Calculating variance decomposition
sigma_e2 = var(y - y_i_bar)            # Within-individual variance
sigma_alpha2 = var(y_i_bar - y_bar)    # Between-individual variance

# Calculating lambda
T = length(unique(time))  # Number of time periods per individual
lambda = 1 - sqrt(sigma_e2 / (sigma_e2 + T * sigma_alpha2))

# Applying transformation
y_trans = y - lambda * y_i_bar
X_trans = X - lambda * X_i_bar

# Estimating random effects model with OLS
re_model = lm(y_trans ~ X_trans)

## Estimating the same model using plm package as benchmark
plm_re = plm(lwage ~ edu + ms + exp + exp_sqrd,
           data = psid_df,
           model = "random",
           index = c("id", "year"))

## Extracting results
coef_random = coef(summary(re_model))[3:5, ]
coef_re = coef(summary(plm_re))[3:5, ]

## Creating a comparison DataFrame
comparison_re = data.frame(
  Estimate_random = coef_random[, "Estimate"],
  Estimate_plm_re = coef_re[, "Estimate"],
  StdError_random = as.double(sprintf("%.10f", coef_random[, "Std. Error"])),
  StdError_plm_re = as.double(sprintf("%.10f", coef_re[, "Std. Error"]))
)
rownames(comparison_re) = c("ms", "exp", "exp_sqrd")

## Summarizing results
comparison_re
```
The main difference between this two models is that in the random effect we were able to estimate the effect of education on the log wage. In terms of point estimation of the covariates, the experience variables shows little difference, either between models and between the benchmark using *plm* package. However, for the marital status variable, we have a significant divergence in the "by-hand" estimation, both in terms of signal and level. In other words, while the benchmark estimation gives a coefficient of -0.0127, the "by-hand" estimation gives a value of 0.0068.

## Item (c)
Calculating the Hausman statistic and performing Hausman test by brute force:
```{r}
coef_fe = coef(within_est)[2:4]
coef_re = coef(re_model)[3:5]
vcov_fe = vcov(within_est)[2:4, 2:4]
vcov_re = vcov(re_model)[3:5, 3:5]
coef_diff = coef_re - coef_fe
vcov_diff = vcov_fe - vcov_re
H = if (as.numeric(t(coef_diff) %*% solve(vcov_diff) %*% coef_diff) > 0) {
  as.numeric(t(coef_diff) %*% solve(vcov_diff) %*% coef_diff)
} else {
  (as.numeric(t(coef_diff) %*% solve(vcov_diff) %*% coef_diff))*(-1)
}
p_value = pchisq(H, df = length(coef_fe), lower.tail = FALSE)
cat("Test statistic:", H, "\nValor-p:", p_value, "\n")
if (p_value < 0.05) {
  cat("Reject H0: Fixed effect model is more appropriate.\n")
} else {
  cat("Do not reject H0: Random effect model is more appropriate.\n")
}
```
Hence, we reject de null hypothesis that there is no correlation between individual-specific unobserved heterogeneity and individuals’ characteristics captured by the covariates (marital status, experience and squared experience), which means that the fixed effect model is more appropriate. Recall that we can not compare the models in terms of the education regressor, since it was suppressed in the fixed effect estimate.
Finally, notice that estimated variances for the entity-specific time-invariant unobserved heterogeneity and for the idiosyncratic component are represented by *sigma_e2* and *sigma_alpha2* objects, respectively.
